---
title: "Data Cleaning In R"
author: "Waqas"
date: "2022-09-27"
output: 
  html_document:
    toc: yes
    theme: united
---
# Importing packages for data cleaning

```{r = importing packages}
library(tidyverse)
library(lubridate)
library(janitor)
```

* The **tidyverse** package is an overall packages which includes almost everything related to data analysis.

* The **lubridate** package deals with everything related to date.

* The **janitor** package deals with things related to cleaning data.

# Importing Raw Dataset

```{r =Importing raw dataset}
library(palmerpenguins)
data("penguins")
view(penguins_raw)
```

#Checking the datatype of each column

```{r = checking datatype}
str(penguins_raw)
```
# Cleaning the Data
## Separating the species column

```{r = separating the column}
penguin_clean <- penguins_raw %>% separate(Species, c("species", "animal", "family"))
colnames(penguin_clean)
```
This separates the column into three new column,and it separates the data inside them accordingly

## Changing the date format

```{r = changing date format}
penguin_clean$date <- format(as.Date(penguin_clean$`Date Egg`), "%Y")
colnames(penguin_clean)
```

We can see that a new date column is added here, which contains year of the data collection

## Renaming Column Names

```{r = renaming column}
penguin_clean <- penguin_clean %>% rename(bill_length_mm = `Culmen Length (mm)`) %>% rename(bill_depth_mm = `Culmen Depth (mm)`)
colnames(penguin_clean)
```
Here, we renamed some columns to make them more understandable

## Selecting Required Columns

```{r = selecting required columns}
penguin_clean <- penguin_clean %>% select(species, Island, date, Sex, bill_length_mm, bill_depth_mm, `Flipper Length (mm)`, `Body Mass (g)`)
colnames(penguin_clean)
```
This selects only the required columns for analysis purposes

## Making the column names even throughout

```{r = changing column names}
penguin_clean <- penguin_clean %>% clean_names()
colnames(penguin_clean)
```
This converts the column name to lower case and space is replaced by an underscore. This makes the column names look consistent throughout.

## Dropping Empty Rows

```{r = dropping empty rows}
penguin_clean <- penguin_clean %>% drop_na()
```
This removes all the empty rows, thus makes the data more useful and clean.

Thus, our data is clean now and ready for analysis





